---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# findBiobankR

<!-- badges: start -->

 [![R-CMD-check](https://github.com/finddx/findBiobankR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/finddx/findBiobankR/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

findBiobankR package is a collection of functions used by FindDx Biobank 

## Installation

You can install the development version of findBiobankR from [GitHub](https://github.com/) with:

``` r

# install.packages("devtools")
devtools::install_github("finddx/findBiobankR")

```

## Convert Open Specimen Data to wide for Selection 

- The function below is used to prepare clinical and specimen data sets for sample selections 

```{r convert_os_df_to_wide}
library(findBiobankR)
library(dplyr)
library(data.table)
library(knitr)

data("tb_data", package = "findBiobankR")
data("tb_specimen_df", package = "findBiobankR")

# required clinical cols
clinical_cols <- c("tb_group", "hiv_status", "sex", "country") 

clin_specs  <- convert_os_df_to_wide(clinical_df = tb_data,
                                     specimen_df = tb_specimen_df, 
                                     join_by_col = "ppid",
                                     specimen_col ="specimen_type" ,
                                     clinical_vars = clinical_cols ,
                                     specimen_label_col = "specimen_label",
                                     os_aliquot_names = c("Serum", "Plasma")
)

knitr::kable(clin_specs[1:5, ])
```


## Select Specimen

- Using the data set above select aliquot 

```{r}
##
samples_list <- list(c("TB0990100010101001", "TB0990100010101002", "TB0990100010101003"),
                     c("TB0990100020101001", "TB0990100020101002", "TB0990100020101003"), 
                     c("TB0990100030101001", "TB0990100030101002", "TB0990100030101003"),
                     c("TB0990100040101001", "TB0990100040101002", "TB0990100040101003"),
                     c("TB0990100050101001", "TB0990100050101002", "TB0990100050101003"
                     ))
# here aliquot number means index or position of the aliquot 
select_one_aliquot(samples_list,
                   aliquot_number = 1)

# You can use the data table method to already append this to the data set
clin_specs[, serum_aliquot_dt := select_one_aliquot(NP_labels_serum)]
# or tidyverse way
clin_specs <- clin_specs %>%
  mutate(serum_aliquot_dp = select_one_aliquot(NP_labels_serum))

knitr::kable(clin_specs[1:5, ])
```

## Select many aliquots at once

- Let's say a requester wants 2 aliquots per patient. Using the above function it's not convenient


```{r}
##
clin_specs  <- convert_os_df_to_wide(clinical_df = tb_data,
                                     specimen_df = tb_specimen_df, 
                                     join_by_col = "ppid",
                                     specimen_col ="specimen_type" ,
                                     clinical_vars = clinical_cols ,
                                     specimen_label_col = "specimen_label",
                                     os_aliquot_names = c("Serum", "Plasma")
)
# here aliquots stands for number of samples per patient
# so this will select firts and second specimen for each participants
select_many_aliquots(samples_list,
                   aliquots = 2)

# You can use the data table method to already append this to the data set
col_nms <- paste0("aliquot", 1:2, "_dt")
clin_specs[, (col_nms ) := select_many_aliquots(NP_labels_serum, 
                                                aliquots = 2)]

knitr::kable(head(clin_specs, 5))
# unfortunately no easy for now tidyverse implementation but trying this below could work

df <- select_many_aliquots(clin_specs$NP_labels_serum, aliquots = 2) %>% 
  as.data.frame()
# You can then bind cols to the original data
knitr::kable(head(df, 2))
```


## Find is a patient has a positive/negative outcome based on several tests 

- Find if a participant had any of respiratory symptoms 

```{r}
data("tb_resp_symptoms")

respirotory_symptoms <- c("COUGH", "EXPECTORATION",
                          "HEMOPTYSIS", "CHEST_PAIN",
                          "DYSPNOE")
kable(head(tb_resp_symptoms, 5))
outcome_df  <- any_pos_any_neg(df = tb_resp_symptoms, 
                 test_cols =respirotory_symptoms ,
                 test_type = "resp_symp" ,
                 neg_value ="No",
                 pos_value = "Yes",
                 id_cols ="ID")
kable(head(outcome_df, 5))
```

## Select number of patients 

- From `convert_os_df_to_wide` output select 1 patients tb neg and 2 tb pos

```{r}
## select 2 patients per group 
patient_groups  <- c("TB, Neg", "TB, Pos")
number_per_group  <- c(1, 2)
sort_cols = "NP_aliqN_serum"
clin_specs_2 <- select_patients_per_group(df = clin_specs, 
                                          patient_group_col = "tb_group",
                                          patient_groups = patient_groups,
                                          number_per_group = number_per_group,
                                          sort_cols = sort_cols)
kable(clin_specs_2)
```

