---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# findBiobankR

<!-- badges: start -->

 [![R-CMD-check](https://github.com/finddx/findBiobankR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/finddx/findBiobankR/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

findBiobankR package is a collection of functions used by FindDx Biobank 

## Installation

You can install the development version of findBiobankR from [GitHub](https://github.com/) with:

``` r

# install.packages("devtools")
devtools::install_github("finddx/findBiobankR")

```

## Convert Open Specimen Data to wide for Selection 

- The function below is used to prepare clinical and specimen data sets for sample selections 

```{r convert_os_df_to_wide}
library(findBiobankR)
library(dplyr)
library(data.table)
library(knitr)

data("tb_data", package = "findBiobankR")
data("tb_specimen_df", package = "findBiobankR")

# required clinical cols
clinical_cols <- c("tb_group", "hiv_status", "sex", "country") 

clin_specs  <- convert_os_df_to_wide(clinical_df = tb_data,
                                     specimen_df = tb_specimen_df, 
                                     join_by_col = "ppid",
                                     specimen_col ="specimen_type" ,
                                     clinical_vars = clinical_cols ,
                                     specimen_label_col = "specimen_label",
                                     os_aliquot_names = c("Serum", "Plasma")
)

knitr::kable(clin_specs[1:5, ])
```


## Select Specimen

- Using the data set above select aliquot 

```{r}
##
samples_list <- list(c("TB0990100010101001", "TB0990100010101002", "TB0990100010101003"),
                     c("TB0990100020101001", "TB0990100020101002", "TB0990100020101003"), 
                     c("TB0990100030101001", "TB0990100030101002", "TB0990100030101003"),
                     c("TB0990100040101001", "TB0990100040101002", "TB0990100040101003"),
                     c("TB0990100050101001", "TB0990100050101002", "TB0990100050101003"
                     ))
# here aliquot number means index or position of the aliquot 
select_one_aliquot(samples_list,
                   aliquot_number = 1)

# You can use the data table method to already append this to the data set
clin_specs[, serum_aliquot_dt := select_one_aliquot(NP_labels_serum)]
# or tidyverse way
clin_specs <- clin_specs %>%
  mutate(serum_aliquot_dp = select_one_aliquot(NP_labels_serum))

knitr::kable(clin_specs[1:5, ])
```

## Select many aliquots at once

- Let's say a requester wants 2 aliquots per patient. Using the above function it's not convenient


```{r}
##
clin_specs  <- convert_os_df_to_wide(clinical_df = tb_data,
                                     specimen_df = tb_specimen_df, 
                                     join_by_col = "ppid",
                                     specimen_col ="specimen_type" ,
                                     clinical_vars = clinical_cols ,
                                     specimen_label_col = "specimen_label",
                                     os_aliquot_names = c("Serum", "Plasma")
)
# here aliquots stands for number of samples per patient
# so this will select firts and second specimen for each participants
select_many_aliquots(samples_list,
                   aliquots = 2)

# You can use the data table method to already append this to the data set
col_nms <- paste0("aliquot", 1:2, "_dt")
clin_specs[, (col_nms ) := select_many_aliquots(NP_labels_serum, 
                                                aliquots = 2)]

knitr::kable(head(clin_specs, 5))
# unfortunately no easy for now tidyverse implementation but trying this below could work

df <- select_many_aliquots(clin_specs$NP_labels_serum, aliquots = 2) %>% 
  as.data.frame()
# You can then bind cols to the original data
knitr::kable(head(df, 2))
```


## Find if a patient has a positive/negative outcome based on several tests 

- Find if a participant had any of respiratory symptoms 

```{r}
data("tb_resp_symptoms")

respirotory_symptoms <- c("COUGH", "EXPECTORATION",
                          "HEMOPTYSIS", "CHEST_PAIN",
                          "DYSPNOE")
kable(head(tb_resp_symptoms, 5))
outcome_df  <- any_pos_any_neg(df = tb_resp_symptoms, 
                 test_cols =respirotory_symptoms ,
                 test_type = "resp_symp" ,
                # neg_value ="No",
                 pos_value = "Yes",
                 id_cols ="ID")
kable(head(outcome_df, 5))
```

## Select number of patients 

- From `convert_os_df_to_wide` output select 1 patients tb neg and 2 tb pos

```{r}
## select 2 patients per group 
patient_groups  <- c("TB, Neg", "TB, Pos")
number_per_group  <- c(1, 2)
sort_cols = "NP_aliqN_serum"
clin_specs_2 <- select_patients_per_group(df = clin_specs, 
                                          patient_group_col = "tb_group",
                                          patient_groups = patient_groups,
                                          number_per_group = number_per_group,
                                          sort_cols = sort_cols)
kable(clin_specs_2)
```

## Read multiple csv files

- Sometimes you have a folder with a couple of files you want to read

```{r, results='asis'}
library(data.table)
iris_split <- split(iris, 
                    f = iris$Species)


path_name <- tempfile(pattern = "read_multipe_files")
#> Warning in normalizePath(path.expand(path), winslash, mustWork):
#> path[1]="C:\Users\MOSES~1.MBU\AppData\Local\Temp\RtmpAPcQpP\read_multipe_files15d0134c973":
#> The system cannot find the file specified

dir.create(path_name)


path_name <- normalizePath(path_name, winslash = "/")

file_names <- paste0(names(iris_split),  ".csv")
# create file name
file_paths <- file.path(path_name, file_names)
lapply(seq_along(file_paths), function(x){
    df = iris_split[[x]]
    fwrite(df, file = file_paths[x])
})
#> [[1]]
#> NULL
#> 
#> [[2]]
#> NULL
#> 
#> [[3]]
#> NULL

read_multiple_files(folder_path =path_name ,
                    read_function = "fread",
                    pattern = ".csv$",
                    file_names =NULL) %>% 
  lapply( function(x){
    
    kable(head(x, 5))
  })
```

## Get OS Query

```{r}
auth_response <- auth_os(url =  Sys.getenv("OSTESTURL"),
                         username = Sys.getenv("OSUSERNAME"),
                         password = Sys.getenv("OSPASSWORDTEST"))

query114 <- get_os_query(auth_response, query_id = 114,
                         driving_form = "Participant",
                         wide_row_mode = "OFF",
                         start_at = 0,
                         max_results = 5)

kable(query114)

```

## Get OS Bulk Query

```{r}

cv_samples <- get_bulk_query(auth_response, query_id = 105)
  

kable(cv_samples[1:5])
```

## Get os Orders

```{r}

orders <- get_bulk_orders(auth_response)

kable(orders[1:5])
```

## Get Order Details


```{r}
oder_details <- get_order_detail(auth_response, order_id = 52)

kable(oder_details)
```


## Get bulk order details

```{r}
oders_details <- get_bulk_order_detail(auth_response, orders_ids = orders$id )

kable(oders_details[1:5])
```

## Get Shipped/Distributed Samples

```{r}
distributed_samples <- get_bulk_order_items(auth_response,
                                        orders_ids = orders$id )

kable(distributed_samples[1:5])

```

## Get all Sites

```{r}
sites <- get_all_sites(auth_response)
kable(sites[1:5])
```

